Capacity and Utilization analysis of Toronto shelter(2024)

### 2024-09-27 paper/paper_Li.qmd
Update paper_Li.qmd

# install.packages("tidyverse")
# install.packages("opendatatoronto")
# install.packages("here")
library(tidyverse)
library(opendatatoronto)
library(here)
```

@@

@@

```{r}
#| label: fig-occupancy-capital-relat
#| fig-cap: Scatter Plot of Bed Capacity vs. Occupancy Rate
#| echo: false
#| warning: false
#| message: false
# Convert OCCUPANCY_RATE_BEDS to numeric (in case of any issues during import)
analysis_data$OCCUPANCY_RATE_BEDS <- as.numeric(analysis_data$OCCUPANCY_RATE_BEDS)

# Calculate the average occupancy rate per location
avg_occupancy_rate <- aggregate(OCCUPANCY_RATE_BEDS ~ LOCATION_NAME, analysis_data, mean, na.rm = TRUE)

# Sort the data by occupancy rate
avg_occupancy_rate <- avg_occupancy_rate[order(avg_occupancy_rate$OCCUPANCY_RATE_BEDS), ]

# Plot the bar chart with smaller text
ggplot(avg_occupancy_rate, aes(x = reorder(LOCATION_NAME, OCCUPANCY_RATE_BEDS), y = OCCUPANCY_RATE_BEDS)) +
  geom_bar(stat = "identity", fill = "skyblue", width = 0.6) + # Use a more pleasing color
  coord_flip() + # Flip to make it horizontal
  labs(title = "", x = "Location", y = "Occupancy Rate (%)") +
  theme_minimal() + # Clean theme for better aesthetics
  theme(
    plot.title = element_text(size = 10, face = "bold", hjust = 0.5), # Smaller title size
    axis.title = element_text(size = 9), # Smaller axis title size
    axis.text.x = element_text(size = 7), # Smaller size of x-axis text
    axis.text.y = element_text(size = 3.8), # Smaller size of y-axis text (locations)
    panel.grid.major.y = element_blank() # Remove y-axis grid lines for better readability
  ) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) # Ensure bars stay within 0-100% limits
```

```{r}
#| label: fig-occupancy-capital-123
#| fig-cap: Scatter Plot of Bed Capacity vs. Occupancy Rate
#| echo: false
#| warning: false
#| message: false
# Clean the data by removing NA values in CAPACITY_ACTUAL_BED and OCCUPANCY_RATE_BEDS
scatter_data <- na.omit(analysis_data[, c("CAPACITY_ACTUAL_BED", "OCCUPANCY_RATE_BEDS")])

@@

@fig-occupancy-capital-relat displays the average occupancy rate across various shelters. The X-axis represents the occupancy rate percentage (Occupancy Rate %), ranging from 0% to 100%, while the Y-axis lists the shelter locations (Location), showing the names of several shelters. Each horizontal blue bar corresponds to the average occupancy rate of a specific shelter. Most shelters have an occupancy rate close to 100%, indicating that their beds are almost fully or completely occupied. A few shelters have occupancy rates below 75%, showing that they have relatively lower bed utilization.

@@

@fig-occupancy-capital-relat paints a sobering picture of Toronto's shelter system, with most facilities operating at or near full capacity. This level of demand points to a systemic issue that extends beyond shelter availability to broader socioeconomic challenges. The high occupancy rates reflect not just a lack of shelters but underline key drivers of homelessness, such as inadequate affordable housing, mental health challenges, unemployment, and other economic pressures.

@@

The data cleaning process involved filtering out some of the columns from the raw dataset and renaming some of the data entries for clarity and simplicity.

>>>>

# install.packages("tidyverse")
# install.packages("opendatatoronto")
# install.packages("here")
library(opendatatoronto)
library(tidyverse)
library(here)
library(knitr)
```

@@

```{r}
#| label: tbl-sample
#| echo: false
#| warning: false
#| message: false
# Load the data from the specified path
data <- read.csv("C:/Users/EasonLi/Downloads/Toronto-main/Toronto-main/data/analysis_data_clean.csv")

# Display the first 5 rows with all columns
kable(head(data, 5),
  caption = "Table: Sample of cleaned occupancy data",
  col.names = c(
    "Occupancy Date", "Occupancy Rate (Beds)", "Occupancy Rate (Rooms)",
    "Location Name", "Actual Bed Capacity", "Sector"
  )
)
```

@tbl-sample show a sample of cleaned occupancy data, Since the statistical record formats of different shelters are different, each row contains different information marked as NA, but this does not negattive affect the analysis.
\newpage

@@

```{r}
#| label: fig-occupancy-capital-relat
#| fig-cap: Scatter Plot of Bed Capacity vs. Occupancy Rate
#| echo: false
#| warning: false
#| message: false
#|
# 转换 'OCCUPANCY_RATE_BEDS' 为数值型，过滤掉缺失值
analysis_data <- analysis_data %>%
  mutate(OCCUPANCY_RATE_BEDS = as.numeric(OCCUPANCY_RATE_BEDS)) %>%
  filter(!is.na(OCCUPANCY_RATE_BEDS))

# 计算每个位置和 sector 的平均占用率
avg_occupancy_rate <- analysis_data %>%
  group_by(SECTOR, LOCATION_NAME) %>%
  summarize(OCCUPANCY_RATE_BEDS = mean(OCCUPANCY_RATE_BEDS, na.rm = TRUE)) %>%
  arrange(SECTOR, OCCUPANCY_RATE_BEDS)

sector_colors <- c("Men" = "#66c2a5", "Mixed Adult" = "#fc8d62", "Women" = "#8da0cb", "Youth" = "#e78ac3")

ggplot(avg_occupancy_rate, aes(x = reorder(LOCATION_NAME, OCCUPANCY_RATE_BEDS), y = OCCUPANCY_RATE_BEDS, fill = SECTOR)) +
  geom_bar(stat = "identity", width = 0.6) +
  coord_flip() +
  labs(title = "", x = "", y = " (%)") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 0, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 0),
    axis.text.x = element_text(size = 5),
    axis.text.y = element_text(size = 4),
    panel.grid.major.y = element_blank(),
    legend.position = "none"
  ) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 100)) +
  facet_wrap(~SECTOR, scales = "free_y", ncol = 2) +
  scale_fill_manual(values = sector_colors)
```
```{r}
#| label: fig-occupancy-capital-123
#| fig-cap: Scatter Plot of Bed Capacity vs. Occupancy Rate
#| echo: false
#| warning: false
#| message: false
# Clean the data by removing NA values in CAPACITY_ACTUAL_BED and OCCUPANCY_RATE_BEDS
scatter_data <- na.omit(analysis_data[, c("CAPACITY_ACTUAL_BED", "OCCUPANCY_RATE_BEDS")])

@@

@fig-occupancy-capital-relat displays the average occupancy rate across various shelters, categorized by sector: Men, Women, Mixed Adult, and Youth. The X-axis represents the occupancy rate percentage (Occupancy Rate %), ranging from 0% to 100%, while the Y-axis lists the shelter locations (Location), showing the names of several shelters. Each horizontal blue bar corresponds to the average occupancy rate of a specific shelter. Most shelters have an occupancy rate close to 100%, indicating that their beds are almost fully or completely occupied. A few shelters have occupancy rates below 75%, showing that they have relatively lower bed utilization.

@@

@fig-occupancy-capital-relat reveals that across all sectors—Men, Women, Mixed Adults, and Youth. Both show a sobering picture of Toronto's shelter system, with most facilities operating at or near full capacity—this level of demand points to a systemic issue that extends beyond shelter availability to broader socioeconomic challenges. The high occupancy rates reflect not just a lack of shelters but also underline key drivers of homelessness, such as inadequate affordable housing, mental health challenges, unemployment, and other economic pressures.

@@

The data cleaning process involved selecting some columns of useful: occupancy date, bed and room occupancy rates, location name, actual bed capacity, and sector. To improve clarity, the column "CAPACITY_ACTUAL_BED" is renamed to "Actual Bed Capacity".
